Design and Implementation of a Multithreaded Virtual Machine for Executing Linear Logic Programs

### 2014-07-25 language.tex
Fixed.

By applying the rule, we consume \texttt{count-prices(@1)} and derive
the aggregate which consumes all the \texttt{price(@1, P)} linear
facts. These are summed up and \texttt{total(@1,~12)} is derived. LM
provides several aggregate operations, including the \emph{minimum},
\emph{maximum}, \emph{sum} and \emph{count}.

@@

Due to the restrictions on LM rules, nodes are able to run rules
independently without using other node's facts. Node computation
follows a \emph{don't care} or \emph{committed choice} non-determinism
since any node can be picked to run as long as it contains enough
facts to fire a derivation rule.  Facts coming from other nodes will
arrive in order of derivation but may be considered partially and
there is no particular order among the neighborhood. To improve
concurrency, the programmer is encouraged to write rules that take
advantage of the non-deterministic nature of execution.

>>>>

By applying the rule, we consume \texttt{count-prices(@1)} and derive
the aggregate which consumes all the \texttt{price(@1, P)} linear
facts. These are summed up and \texttt{total(@1,~12)} is derived. LM
provides several aggregate operations, including the \emph{minimum},
\emph{maximum}, \emph{sum} and \emph{count}. Note that the \texttt{.} syntax indicates
an empty list of variables and \texttt{1} is borrowed from linear logic
and represents an empty derivation.

Comprehensions and aggregates are logically justified by the underlying
proof system of the language. Our proof system is extended with greatest
fixed points~\cite{Baelde:2012:LGF:2071368.2071370}, which allows us to describe
recursive definitions such as comprehensions or aggregates.

@@

Due to the restrictions on LM rules, nodes are able to run rules
independently without using other node's facts. Node computation
follows a \emph{don't care} or \emph{committed choice} non-determinism
since any node can be picked to run as long as it contains enough
facts to fire a derivation rule.  Facts coming from other nodes will
arrive in order of derivation but may be considered partially and
there is no particular order among the neighborhood. To improve
concurrency, the programmer is encouraged to write rules that take
advantage of the non-deterministic nature of execution since too much
determinism will naturally imply less scalability and more synchronization
when executing the programs.

### 2014-07-25 virtual_machine.tex
Fixed.

A key goal of our design is to keep the threads as busy as possible
and to reduce inter-thread communication. When the VM starts, it reads
the byte-code file and starts all threads. Initially, the VM will
partition the application graph of $N$ nodes into $T$ subgraphs (the
number of threads) and then each thread will work on their own
subgraph. Reduction of communication between nodes in different threads
is achieved by first ordering the node addresses present in the code in such a way that
closer nodes are clustered together and then partitioning them to
threads. During compilation, we take note of predicates that are used
in rules with communication (arguments with type \emph{node}) and
then build a graph of nodes from the program's axioms. The nodes of
the graph are then ordered by using a breadth-first search algorithm
that changes the nodes of addresses to the domain $[0, n[$, where $n$
is the number of nodes. Once the VM starts, we simply partition the
range $[0, n[$ into $T$ subgraphs.

@@

Whenever a new fact is derived through rule derivation, we need to
update the data structures for the corresponding node. This is trivial
if the thread that derived the fact also owns the node. If that is not
the case, then we have to synchronize since multiple threads might be
updating the same node's data structures. We added a lock and a
boolean flag to each node to protect the access to its data
structures. When the flag is activated, it means that the node is
currently being executed by the owning thread. For example, in
Fig.~\ref{fig:overview}, if thread 2 derives a fact to
node \texttt{@1} (owned by thread 1), then thread 2 checks the node's
flag and if not activated, will lock node \texttt{@1} and perform the
required updates (\emph{New fact (1)}). If the flag is activated, it
will not touch the main node data structures, but instead will add the
new fact to \emph{Fact Buffer} (\emph{New fact (2)}). The facts stored
in \emph{Fact Buffer} will then be processed whenever the
corresponding node is about to be executed.

@@

Where $A$ is the target argument, $F$ is the multi-set of linear facts
for the target predicate, $values(F, A)$ is set of values of the
argument $A$, $count(F, A = v)$ counts the number of linear facts
where argument $A$ is equal to $v$ and $total(F)$ counts the number of
linear facts in $F$.  The entropy value is a good metric because it
tells us how much information is needed to describe an argument.  If
more information is needed, then that must be the best argument to
index.

>>>>

A key goal of our design is to keep the threads as busy as possible
and to reduce inter-thread communication. When the VM starts, it reads
the byte-code file and starts all threads. Initially, the VM will
partition the application graph of $N$ nodes into $T$ subgraphs (the
number of threads) and then each thread will work on their own
subgraph. Reduction of communication between nodes in different threads
is achieved by first ordering the node addresses present in the code in such a way that
connected nodes are clustered together and then partitioning them to
threads. During compilation, we take note of predicates that are used
for communication (arguments with type \emph{node}) and
then build a graph of nodes from them. The nodes of
the graph are then ordered by using a breadth-first search algorithm
that changes the nodes of addresses to the domain $[0, n[$, where $n$
is the number of nodes. Once the VM starts, we simply partition the
range $[0, n[$ into $T$ subgraphs.

@@

Whenever a new fact is derived through rule derivation, we need to
update the data structures for the corresponding node. This is trivial
if the thread that derived the fact also owns the node. If that is not
the case, then we have to synchronize since multiple threads might be
updating the same node's data structures. We added a lock and a
boolean flag to each node to protect the access to its data
structures. When the flag is activated, it means that the node is
currently being executed by the owning thread. For example, in
Fig.~\ref{fig:overview}, if thread 2 derives a fact to
node \texttt{@1} (owned by thread 1), then thread 2 checks the node's
flag and if not activated, will lock node \texttt{@1} and perform the
required updates (\emph{New fact (1)}). If the flag is activated, it
will not touch the main node data structures, but instead will add the
new fact to \emph{Fact Buffer} (\emph{New fact (2)}). The facts stored
in \emph{Fact Buffer} will then be processed whenever the
corresponding node's flag becomes active.

@@

Where $A$ is the target argument, $F$ is the set of linear facts
for the target predicate, $values(F, A)$ is set of values of the
argument $A$, $count(F, A = v)$ counts the number of linear facts
where argument $A$ is equal to $v$ and $total(F)$ counts the number of
linear facts in $F$.  The entropy value is a good metric because it
tells us how much information is needed to describe an argument.  If
more information is needed, then that must be the best argument to
index.
